Conservative Model Order Reduction for Fluid Flow

### 2018-07-17 sections/1.intro.tex
all corrections added, a backupfile included

Constructing MOR techniques and RB methods that preserve intrinsic structures has recently attracted attention \cite{doi:10.1137/17M1111991,1705.00498,kalashnikova2014stabilization,farhat2015structure,doi:10.1137/110836742,beattie2011structure,doi:10.1137/140978922}. Structure preservation can recover a physically meaningful reduced model, rather than a pure algebraic coupling of equations. This enforces robustness and can help with the stability of the reduced model. Preserving time-symmetries of Lagrangian, Hamiltonian, and port-Hamiltonian systems can be found in the works of \cite{Carlberg:2014ky,doi:10.1137/140978922,doi:10.1137/17M1111991,1705.00498,beattie2011structure,chaturantabut2016structure,gugercin2012structure}. Conserving inf-sup stability, in the context of finite element methods, can be found in \cite{farhat2015structure,ballarin2015supremizer}. Furthermore, a flux preserving model reduction for finite volume methods is presented in \cite{carlberg2018conservative}.

@@

Skew-symmetric formulation of fluid flows construct a skew-symmetric differential operator, acting on the momentum vector field, that ensures conservation of quadratic invariants, such as energy. Combined wirh centered time and space discretization schemes, typically a finite differences discretization method, they recover time-symmetries of a fluid at the discrete level. Such discretization schemes are studies comprehensively over the past few decades and can be found in the works of \cite{morinishi2010skew,morinishi1998fully,desjardins2008high,reiss2014conservative,reiss2014conservative} and the references therein.

In this paper we discuss how to preserve skew-symmetry of the differential operators at the level of the reduced system. This results in conservation of quadratic invariants. Furthermore, we show that the reduced system, as a system of coupled differential equations, contains quadratic invariants and an associated energy which approximates the energy of the high-fidelity system. Therefore, a proper time stepping scheme preserves the reduced representation of the energy, and therefore, the loss in energy due to model reduction remains constant in time.

>>>>

Constructing MOR techniques and RB methods that preserve intrinsic structures has recently attracted attention \cite{doi:10.1137/17M1111991,kalashnikova2014stabilization,farhat2015structure,doi:10.1137/110836742,beattie2011structure,doi:10.1137/140978922}. Structure preservation can recover a physically meaningful reduced model, rather than a pure algebraic coupling of equations. This enforces robustness and can help with the stability of the reduced model. Preserving time-symmetries of Lagrangian, Hamiltonian, and port-Hamiltonian systems can be found in the works of \cite{Carlberg:2014ky,doi:10.1137/140978922,doi:10.1137/17M1111991,beattie2011structure,chaturantabut2016structure,gugercin2012structure}. Conserving inf-sup stability, in the context of finite element methods, can be found in \cite{farhat2015structure,ballarin2015supremizer}. Furthermore, a flux preserving model reduction for finite volume methods is presented in \cite{carlberg2018conservative}.

@@

Skew-symmetric formulation of fluid flows construct a skew-symmetric differential operator, acting on the momentum vector field, that ensures conservation of quadratic invariants, such as energy. Combined wirh centered time and space discretization schemes, typically a finite differences discretization method, they recover time-symmetries of a fluid at the discrete level. Such discretization schemes are studies comprehensively over the past few decades and can be found in the works of \cite{morinishi2010skew,morinishi1998fully,desjardins2008high,tadmor1984skew,reiss2014conservative,reiss2014conservative} and the references therein.

In this paper we discuss how to preserve skew-symmetry of the differential operators at the level of the reduced system. This results in conservation of quadratic invariants. Furthermore, we show that the reduced system, as a system of coupled differential equations, contains quadratic invariants and an associated energy which approximates the energy of the high-fidelity system. Therefore, a proper time stepping scheme preserves the reduced representation of the energy, and therefore, the loss in energy due to model reduction remains constant in time. Furthermore, we demonstrate, through numerical experiments, that a quasi-skew-symmetric form of fluid flow, i.e. a formulation where only spacial differential operators are in a skew-symmetric form, offer remarkable stability properties in terms of MOR. This allows an explicit time-integration to be utilized while recovering robustness of skew-symmetric forms at the reduced level.

### 2018-07-17 sections/3.skew.tex
all corrections added, a backupfile included

In this section we summarize the conservation properties of skew-symmetric forms and discretization schemes, following, closely, the works of \cite{morinishi2010skew,morinishi1998fully,reiss2014conservative}.

>>>>

In this section we summarize the conservation properties of skew-symmetric forms and discretization schemes, following, closely, the works of \cite{morinishi2010skew,morinishi1998fully,tadmor1984skew,reiss2014conservative}.
